---
title: "Lab 2"
author: "Your Name"
date: ''
output:
  pdf_document: 
    fig_caption: yes
header-includes:
  \usepackage{float}
  \floatplacement{figure}{H}
  \usepackage{hyperref}
  \hypersetup{
    colorlinks=true,    
    urlcolor=cyan
  }
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.pos = 'h')
```

## Probability distributions in R and Stan

# Task 1 (8 points)

Study the previous code provided on probability distribution functions in R (see the \texttt{Prob\_Distribs.RMD} file on BbLearn). Your task is to choose a continuous probability distribution other than the Gaussian distribution and complete the following (2 points each). You can use the Stan functions \href{https://mc-stan.org/docs/2_20/functions-reference/index.html#overview}{documentation} to help.   

\begin{enumerate}
  \item Create a custom function to calculate the relative probability density of a given observation of a continuous random variate.
  \item Compare your function to R's built-in density function (i.e., show that they give the same output).
  \item Plot the probability density function $P(y | \Theta)$, where $\Theta$ is a vector of user-specified parameters that define the probability distribution of interest. 
  \item Plot the cumulative density function $F(y | \Theta)$, as above. 
\end{enumerate}

\newpage
# Task 2 (12 points)

```{r required, message=FALSE}
# Load the rstan package
library(rstan)
# Set some useful options
options(mc.cores = parallel::detectCores())
rstan_options(auto_write = TRUE)
```

Our next goal is to learn more about Stan's syntax for probability distributions, which will be critical for defining prior probabilities for our model parameters, as well as defining model likelihoods. 

Below I demonstrate how to use Stan to understand the Gaussian (Normal) distribution. I will simulate random draws and calculate relative probability densities.

```{r stan-sim}

# Define data inputs

## Number of random draws:
n_draws = 1000
## Number of points used for PDF:
n_seq = 500
## Sequence of observed y, for PDF:
y_pdf = seq(-8, 10, length.out = n_seq)

# Store the required data in a list
sim_data = list(n_draws = n_draws,
                n_seq = n_seq,
                y_pdf = y_pdf)

#------------------
#------------------

# Run the simulation:

sim_fit =
  stan(file="pdf_sim.stan",
       data=sim_data,
       iter=1,
       chains=1,
       algorithm="Fixed_param")

# The function extract() will create a list of output
# Note that "lp__" is irrelevant here, but will become
# very important later in the course. 

sim_out = extract(sim_fit)
str(sim_out)

# Plot the random draws:
hist(sim_out$rand_norm[, 1:n_draws], main = "", xlab = "y")
mean(sim_out$rand_norm[, 1:n_draws])
var(sim_out$rand_norm[, 1:n_draws])

# Plot the PDF

## First get back to usual scale:
lik_pdf = exp(sim_out$log_lik[, 1:n_seq])

## Now plot:
plot(lik_pdf ~ y_pdf, type = "l",
     xlab = "y", ylab = expression("P(y | "~mu~","~sigma^2~")"))

# Compare to R's PDF function:

lik_pdf_R = dnorm(y_pdf, mean = 1.0, sd = sqrt(4.0))

plot(lik_pdf_R ~ y_pdf, type = "l", col = "red",
     xlab = "y", ylab = expression("P(y | "~mu~","~sigma^2~")"))

```


Your task is to now choose a \textbf{discrete} probability distribution, and use the guided code above to complete the following tasks. Again, see the Stan functions \href{https://mc-stan.org/docs/2_20/functions-reference/index.html#overview}{documentation} for assistance.

\begin{enumerate}
  \item Create a \texttt{pmf\_sim.stan} file that generates the PMF of your chosen distribution and that draws random values from the distribution (5 points).
  \item Plot a histogram of the random draws (2 points).
  \item Plot the PMF using bars or points (3 points).
  \item Compare the PMF generated in Stan to the PMF generated by R's built-in functions (2 points). 
\end{enumerate}






